\modLP{Identification of} Biological Regulatory Networks from Process Hitting models

### 2014-11-03 parts-asp.tex
end switching from interval parameters to integer parameters

\begin{example}\label{ex:asp-atom}
Consider a PH model such as \pref{ex:infer-param-runningPH-1}:
in order to state the existence of each component, we use a predicate called $\component$ with two arguments:
\begin{align*}
  &\component(x, n).
\end{align*}
where $x$ is the name of the component and $l_x = n$ is its maximum expression level.
\end{example}

@@

\begin{example}\label{ex:asp-component}
In order to define the 3 components of \pref{ex:infer-param-runningPH-1}, we use the following program:
\begin{align*}
  &\component(a, 2). \\
  &\component(b, 1). \\
  &\component(c, 1).
\end{align*}
This program contains only facts using the predicate $\component$ defined in \pref{ex:asp-atom},
and $a$, $b$, $c$, $1$ and $2$ are constants.
\end{example}

@@

\begin{example}
Regarding \pref{ex:infer-param-runningPH-1}, the previous rule together with the facts of \pref{ex:asp-component}
will give the following answer set:
\begin{align*}
  &\{&&\component(a, 2),
  &&\component(b, 1), \\
  &&&\component(c, 1),
  &&\componentlevels(a, 0), \\
  &&&\componentlevels(a, 1),
  &&\componentlevels(a, 2), \\
  &&&\componentlevels(b, 0),
  &&\componentlevels(b, 1), \\
  &&&\componentlevels(c, 0),
  &&\componentlevels(c, 1). &\}
\end{align*}
\end{example}

@@

\begin{example}\label{ex:cardinality}
\todo{Refresh this example with integer parameters}
In the scope of \pref{ex:infer-param-runningPH-1}, $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ could not be inferred by \pref{pps:param_K}.
The previous cardinality rule allows to produce 49 parametrizations, in which these two parameters can take all possible values:
%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
%(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
\begin{align*}
  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
\end{align*}
and all the other parameters keep their inferred values.
%These parametrizations take the form of sets of $\param$ atoms.
Note that $\{ 0, 2 \}$ belongs to the set of candidate parametrizations
as no rule specifying that a parameter has to be an interval has been defined yet.
\end{example}

@@

\begin{example}
\todo{Refresh this example with integer parameters}
The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
For example, a constraint was written in order to obtain only interval parameters
(thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
\begin{align*}
  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
\end{align*}

Then, due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
the monotonicity assumption (\pref{pro:param_enum_monotonicity} and expressed in ASP above) removes all parametrizations with:
$0 \in K_{a,\{a,b\}} \vee 0 \in K_{a,\{a,c\}}$.
In the end, the remaining possible values for the two parameters that were not inferred are:
\begin{align*}
  (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{1}{1} , \segm{2}{2} , \segm{1}{2} \}^2
\end{align*}
All these candidates respect the other properties of \pref{ssec:admissible-K} and no more candidates are filtered.
We thus obtain the results of \pref{ex:enum-param-runningPH-1}.
\end{example}

>>>>

\begin{example}\label{ex:asp-atom}
Consider a PH model such as depicted in \pref{fig:runningPH-1}:
in order to state the existence of each component, we use a predicate called $\component$ with two arguments:
\begin{align*}
  &\component(x, n).
\end{align*}
where $x$ is the name of the component and $l_x = n$ is its maximum expression level.
\end{example}

@@

\begin{example}\label{ex:asp-component}
In order to define the 3 components of \pref{fig:runningPH-1}, we use the following program:
\begin{align*}
  &\component(a, 2). \\
  &\component(b, 1). \\
  &\component(c, 1).
\end{align*}
This program contains only facts using the predicate $\component$ defined in \pref{ex:asp-atom},
and $a$, $b$, $c$, $1$ and $2$ are constants.
\end{example}

@@

\begin{example}
Regarding \pref{fig:runningPH-1}, the previous rule together with the facts of \pref{ex:asp-component}
will give the following answer set:
\begin{align*}
  &\{&&\component(a, 2),
  &&\component(b, 1), \\
  &&&\component(c, 1),
  &&\componentlevels(a, 0), \\
  &&&\componentlevels(a, 1),
  &&\componentlevels(a, 2), \\
  &&&\componentlevels(b, 0),
  &&\componentlevels(b, 1), \\
  &&&\componentlevels(c, 0),
  &&\componentlevels(c, 1). &\}
\end{align*}
\end{example}

@@

\begin{example}\label{ex:cardinality}
\modMF{%
In the scope of \pref{ex:infer-param-runningPH-1},
$K_{a,\{b\}}$, $K_{a,\{c\}}$ and $K_{b,\emptyset}$
could not be inferred by \pref{pps:param_K}.
The previous cardinality rule allows to produce 18 parametrizations,
in which these three parameters can take all possible values:
%The enumeration thus produces 36 parametrizations, in which these parameters can take all possible values:
%(as, in this case, the assumptions of \pref{ssec:admissible-K} bring no new constraint):
\begin{align*}
  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 0, 1 \}
%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2}, \{ 0, 2 \} \}^2
\end{align*}
and all the other parameters keep their inferred values.
Note that $0$ belongs to the set of candidate values for parameter $K_{b,\emptyset}$
because the extreme values assumption (\pref{pro:param_enum_extreme})
has not been taken into account yet.
%These parametrizations take the form of sets of $\param$ atoms.
% Note that $\{ 0, 2 \}$ belongs to the set of candidate parametrizations
% as no rule specifying that a parameter has to be an interval has been defined yet.
}%
\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
\end{example}

@@

\begin{example}
\modMF{%
The set of candidate values given in \pref{ex:cardinality} can be filtered using constraints.
% For example, a constraint was written in order to obtain only interval parameters
% (thus avoiding parameters with “gaps” such as $\{ 0, 2 \}$).
% Applying such a constraint would reduce the set of possible values for the parameters $K_{a,\{a,b\}}$ and $K_{a,\{a,c\}}$ to:
% \begin{align*}
%   (K_{a,\{a,b\}} ; K_{a,\{a,c\}}) \in \{ \segm{0}{0} , \segm{1}{1} , \segm{2}{2} , \segm{0}{1} , \segm{1}{2} , \segm{0}{2} \}^2
% \end{align*}
% }%
% 
% \modMF{%
For example,
% , due to the fact that the parameters $K_{a,\{b\}} = \segm{1}{1}$ and $K_{a,\{c\}} = \segm{1}{1}$ could be inferred,
the extreme values assumption (\pref{pro:param_enum_extreme}) removes all parametrizations with
$K_{b,\emptyset} = 0$,
thus reducing the set of candidate values to 9 possible values:
\begin{align*}
  (K_{a,\{b\}}; K_{a,\{c\}}; K_{b,\emptyset}) \in
    \{ 0, 1, 2 \} \times \{ 0, 1, 2 \} \times \{ 1 \}
\end{align*}
Furthermore, parameter $K_{b,\emptyset} = 1$ doest not respect \pref{pro:K-valid},
because no action in the model of \pref{fig:runningPH-1} allows to bounce from $b_0$ to $b_1$.
Therefore, this parameter is removed from all returned parametrizations.
In the end, 9 partial parametrizations are therefore produced.
}%
\todo{Find an example to apply the monotonicity assumption?}
\todo{Check this example w.r.t.~\pref{pro:K-valid}.}
\end{example}

### 2014-11-03 parts-example.tex
end switching from interval parameters to integer parameters

The inference methods described in this paper have been implemented as part of
\textsc{Pint}\footnote{Available at \url{http://loicpauleve.name/pint}}, which gathers PH related tools.
Our implementation mainly consists in ASP programs that are solved using
Clingo\footnote{Available at \url{http://potassco.sourceforge.net}}.
The IG and parameter inferences can be performed using the following command:\\
  \hspace*{\parindent}\texttt{ph2thomas -i model.ph -{}-dot ig.dot}\\
where \texttt{model.ph} is the PH model file in \textsc{Pint} format,
and \texttt{ig.dot} is an output file to write the inferred IG in DOT format.
The (possibly partial) inferred parametrization will be returned on the standard output.
The admissible parametrizations enumeration is performed when adding the \texttt{-{}-full-enumerate} parameter to the command,
or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
\todo{Remove \texttt{full-enumerate}?}

@@

\subsection{Biological application}\label{ssec:appli}

>>>>

The inference \modMF{and enumeration} methods described in this paper have been implemented as part of
\textsc{Pint}\footnote{Available at \url{http://loicpauleve.name/pint}}, which gathers PH related tools.
Our implementation mainly consists in ASP programs that are solved using
Clingo\footnote{Available at \url{http://potassco.sourceforge.net}}.
The IG and parameter inferences can be performed using the following command:\\
  \hspace*{\parindent}\texttt{ph2thomas -i model.ph -{}-dot ig.dot}\\
where \texttt{model.ph} is the PH model file in \textsc{Pint} format,
and \texttt{ig.dot} is an output file to write the inferred IG in DOT format.
The (possibly partial) inferred parametrization will be returned on the standard output.
\modMF{%
Instead of the \texttt{-{}-dot ig.dot} option, it is possible to specify
an alternative IG to perform the parameters inference on
(instead of using the inferred IG) with the \texttt{-{}-ig model.ig} option,
where \texttt{model.ig} is an IG model file in \textsc{Pint} format.
The admissible parametrizations enumeration is performed by adding the \texttt{-{}-enumerate}
option to the command.
%or \texttt{-{}-enumerate} instead to obtain only simple parameters (intervals of cardinality $1$).
}%

@@

\subsection{Biological application: the epithelial growth factor receptor}\label{ssec:appli}
